---
title: "Course5-Assignment1"
author: "Vikram"
date: "10/1/2020"
output: html_document
---

#** Physical Activity Monitoring Assignment**
## *Reproducible Research-Week2-Assignment*##

-To begin with INSTALL and LOAD the follwoing packages.
     (dplyr, ggplot2, lubridate, knitr, markdown).
```{r Loading Packages, echo=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)
library(knitr)
library(markdown)

##Objective 1: Load and pre-process the data. 

```{r Load_Data, echo=TRUE}
Data_assignment<-read.csv(file= "activity.csv", header = TRUE, sep=",")

str(Data_assignment)
head(Data_assignment)
tail(Data_assignment)
```


###Data Pre-Processing with column of dates.###

```{r Data Preprocessing, echo=TRUE}
if (FALSE) {
  ymd<- NULL
     day <- NULL
 month <- NULL
 select<- NULL
 
}
Data_assignment$Dateform<- ymd(Data_assignment$date)
Data_assignment$Day<-day(Data_assignment$Dateform)
Data_assignment$Month<-month(Data_assignment$Dateform)
Data_assignment$weekend <- as.factor(ifelse(weekdays(Data_assignment$Dateform)=="Saturday" | weekdays(Data_assignment$Dateform)=="Sunday","weekend","weekday"))
Data_assignment$dayofweek <- as.factor(weekdays(Data_assignment$Dateform))

head(Data_assignment)
tail(Data_assignment)
str(Data_assignment)
```



##Objective 2: Calculate mean total number of steps taken per day by creating Histogram of the total number of steps taken each day.##

```{r Mean Total Number of Steps, echo=TRUE}
Steps_Day <- Data_assignment%>% group_by(Dateform) %>% summarise(Steps_Day = sum(steps,na.rm = TRUE))
str(Steps_Day)
head(Steps_Day)
tail(Steps_Day)
Plot1<-qplot(Steps_Day,data=Steps_Day,na.rm=TRUE,binwidth=1000,
xlab= "Total Steps taken per day", 
ylab="Frequency using binwith 1000",main = "Histogram of the Total number of Steps per Day")

Plot1 

dev.copy(png,'Plot1.png')
dev.off()
```



##Objective 3: To calulate the average daily activity pattern?##

###Mean and Median number of steps taken each day.###

```{r, Average Activity, echo=TRUE}
Steps_Day_mean_median<- Steps_Day%>% summarise(average_steps = mean(Steps_Day,na.rm = TRUE),
median_steps=median(Steps_Day,na.rm = TRUE))

Steps_Day_mean_median
```


###Time series plot of the average number of steps taken.###

```{r Time series graph, echo=TRUE}
TimeSeries_Steps<- Data_assignment %>%  group_by(interval) %>% 
summarize(Steps_mean=mean(steps, na.rm=TRUE))

head(TimeSeries_Steps)
tail(TimeSeries_Steps)
str(TimeSeries_Steps)

Plot2<-ggplot(TimeSeries_Steps, aes(interval, Steps_mean)) + geom_line()

Plot2

dev.copy(png, 'Plot2.png')
dev.off()
```


###Calculate the 5-minute interval that contains the maximum number of steps?###

```{r Interval showing maximum number of steps, echo=TRUE}
Int_max_steps<- TimeSeries_Steps[TimeSeries_Steps$Steps_mean==max(TimeSeries_Steps$Steps_mean),]
Int_max_steps
```



##Objective 4:Imputing missing values.

###Calculate and report the total number of missing values in the dataset.###

```{r Number of missing values, echo=TRUE}
Data_without_NA <- na.omit(Data_assignment)
Rows_NA<- nrow(Data_assignment)-nrow(Data_without_NA)
Rows_NA
```


###Impute missing values by replacing with the mean of the number of steps for the 5-minute interval across all the days and generate the dataset with imputed values.###

```{r Filling missing values, echo=TRUE}
names(TimeSeries_Steps)[2] <- "Mean_Steps"
Dataset_imputed <- merge(Data_assignment, Data_without_NA)

Dataset_imputed$steps[is.na(Dataset_imputed$steps)] <- Dataset_imputed$Mean_Steps[is.na(Dataset_imputed$Mean_Steps)]

summary(Dataset_imputed)
str(Dataset_imputed)
head(Dataset_imputed)
tail(Dataset_imputed)
```


###Create Histogram for total number of steps taken each day from the imputed data.###

```{r Histogram with imputed data, echo=TRUE}
SumSteps_Day_newdata <- Dataset_imputed %>%  group_by(Dateform) %>% 
summarize(Total_steps_day=sum(steps))

summary(SumSteps_Day_newdata)
str(SumSteps_Day_newdata)

Plot3<-hist(SumSteps_Day_newdata$Total_steps_day, xlab = "Steps", ylab = "Frequency", main = "Total Daily Steps", breaks=50)
Plot3

dev.copy(png, 'Plot3.png')
dev.off()
```


###Compare the mean and median of Old and New data.###

```{r Comparing data for the old and the new dataset, echo=TRUE}
NewData<-summary(Dataset_imputed$steps)
OldData<-summary(Data_assignment$steps, na.rm=TRUE)

NewData
OldData
```



##Objective 7: Are there differences in activity patterns between weekdays and weekends?
*Dataset_imputed already has been created using the Dataset_assignment data
which has been converted to the appropriate columns of Date class and the weekday and weekend factors included.*

```{r Dataset_imputed with weekend/ay columns, echo=TRUE}
str(Dataset_imputed)
```


###Create a time series plot of the 5-minute interval (x-axis) and the average number of steps taken,
averaged across all weekday days or weekend days (y-axis).###

```{r Time series plot, echo=TRUE}
Average_Steps<- Dataset_imputed %>% group_by(interval,weekend) %>%   
  summarise(Average_Steps = mean(steps))

summary(Average_Steps)

Plot4<-ggplot(Average_Steps, aes(x=interval, y=Average_Steps, color=weekend)) + geom_line(linetype=1)+ facet_grid(weekend~.)+ xlab("Interval") + ylab("Mean_Steps") +
ggtitle("Comparison of Average Number of Steps")

Plot4

dev.copy(png, "Plot4.png")
dev.off()
```


===========================================================================================




